---
title: "Pset Prediction Assignment Writeup"
author: "Roman Abashin"
date: "3/31/2019"
output: html_document
---

# Loading data and libraries

```{r Setup, comment=FALSE, message = FALSE, echo=FALSE, warning=FALSE}
# Load data
test <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
                 stringsAsFactors = F)
train <- read.csv("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
                  stringsAsFactors = F)

# Load libraries
library(caret)
library(e1071)
library(ggplot2)
library(plyr)
library(tidyverse)
```


This assignment is about building a robust model to predict physical activity type from data collected by a wearable activity tracker like a Fitbit. 

# Explorational Data Analysis
## Target Variable
Our target variable, `classe`, is a 5-level factor variable that indicates the type of activity a person is doing.

```{r Target variable}
qplot(train$classe)
```

Running `qplot(train$classe)` (above) or `table(t(colwise(class)(train)))` shows two things:

* First, that there are no missing values.
* Second, that we have a slight imbalance in our training data. 

The latter will be ignored for our first model, but can definitely be a point for further improvement.

Furthermore, `class(data$classe)` shows that our target variable is encoded as a factor. This will be needed to be dealt with.

## Features
Running `table(t(colwise(class)(train)))` shows, that we 


# Data cleaning

## Encode target variable as factor

## Encode 